Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models

### 2020-06-05 analysis/paper/paper.Rmd
added a paragraph to Conclusions section. updated several figures.

---
title: "Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@biomark.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@biomark.com
      institute: [biomark]
      correspondence: false
  - Sarah L. Hoffmann:
      email: Sarah.Hoffmann@biomark.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@biomark.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
# csl: "../templates/canadian-journal-of-fisheries-and-aquatic-sciences.csl"
abstract: |
  Needs more...

@@

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

@@

The results of the QRF parr capacity model for spring/summer Chinook salmon meet many biological expectations. Focusing on the partial dependence plots (Figure \@ref(fig:pdp-figure)), the QRF model predicts capacity to increase when the wetted width, discharge and the width-to-depth ratio grow, when temperatures are cooler, when there is less fine sediment, when there is more fish cover, when channel unit frequency increases and when the standard deviation of the wetted depth (a proxy for complexity) increases. These are all patterns that emerged from the fish-habitat data, and those fish-habitat relationships match what has been found qualitatively in many other studies (**need citations??**) [@Mossop2006].

@@

@@

```{r rel-imp-figure, fig.cap = "Relative importance of each habitat covariate included in the quantile regression forest (QRF) model to predict habitat capacity, during summer months, for spring/summer Chinook salmon parr"}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  arrange(desc(imp)) %>%
  left_join(hab_dict %>%
              filter(is.na(MetricGroupName) | MetricGroupName != 'Tier 1 Summary')) %>%
  mutate_at(vars(ShortName, Name),
            list(~ fct_reorder(., .x = imp))) %>%
  ggplot(aes(x = Name,
             y = imp)) +
  # geom_col(aes(fill = MetricCategory)) %>%
  geom_col(fill = 'gray40') +
  coord_flip() +
  labs(x = 'Metric',
       y = 'Relative Importance')

@@

```{r site-photo-high, fig.cap = 'Photo of CHaMP site, CBW05583-415923, with predicted high capacity.'}
knitr::include_graphics('../figures/SitePhoto_CBW05583-415923.jpg')
```

```{r site-photo-low, fig.cap = 'Photo of CHaMP site, CBW05583-109994, with predicted low capacity.'}
knitr::include_graphics('../figures/SitePhoto_CBW05583-109994.jpg')
```

>>>>

---
title: "Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models"
author:
  - Kevin E. See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
  - Michael W. Ackerman:
      email: Mike.Ackerman@biomark.com
      institute: [biomark]
      correspondence: false
  - Richard A. Carmichael:
      email: Richie.Carmichael@biomark.com
      institute: [biomark]
      correspondence: false
  - Sarah L. Hoffmann:
      email: Sarah.Hoffmann@biomark.com
      institute: [biomark]
      correspondence: false
  - Chris Beasley:
      email: Chris.Beasley@biomark.com
      institute: [biomark]
      correspondence: false
institute:
  - biomark: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- references.bib
- packages.bib
csl: "../templates/ecological-applications.csl" # Insert path for the bib-style
# csl: "../templates/canadian-journal-of-fisheries-and-aquatic-sciences.csl"
abstract: |
  Needs more...

@@

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 600
)
```

@@

The results of the QRF parr capacity model for spring/summer Chinook salmon meet many biological expectations. Focusing on the partial dependence plots (Figure \@ref(fig:pdp-figure)), the QRF model predicts capacity to increase when the wetted width, discharge and the width-to-depth ratio grow, when temperatures are cooler, when there is less fine sediment, when there is more fish cover, when channel unit frequency increases and when the standard deviation of the wetted depth (a proxy for complexity) increases. These are all patterns that emerged from the fish-habitat data, and those fish-habitat relationships match what has been found qualitatively in many other studies (**need more citations??**) [@Mossop2006].

@@

If a species' carrying capacity is defined or constrained, at least in part, by the habitat in which it lives, then illuminating statistically how such habitat impacts carrying capacity can lead to understanding how a species interacts with its environment. This understanding could be of crucial importance in the realm of conservation, when dealing with an endangered or threatened species, but species/habitat interactions are a core element of ecological studies more generally. We have demonstrated how a quantile regression approach, coupled with a random forest framework, can be used to estimate these relationships, and predict an habitat's capacity. As large ecological datasets become more accessible, and the ability to measure large swaths of habitat more feasible, we believe this approach has many potential applications, from the North American breeding bird survey to groundfish trawl surveys. ***other potential examples (coral reefs? turtles?)***

@@

```{r rel-imp-figure, fig.cap = "Relative importance of each habitat covariate included in the quantile regression forest (QRF) model to predict habitat capacity, during summer months, for spring/summer Chinook salmon parr"}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  arrange(desc(imp)) %>%
  left_join(hab_dict %>%
              filter(is.na(MetricGroupName) | MetricGroupName != 'Tier 1 Summary')) %>%
  mutate_at(vars(ShortName, Name, MetricCategory),
            list(~ fct_reorder(., .x = imp))) %>%
  ggplot(aes(x = Name,
             y = imp)) +
  # geom_col(aes(fill = MetricCategory)) +
  # scale_fill_brewer(palette = 'Set2') +
  geom_col(fill = 'gray40') +
  coord_flip() +
  labs(x = 'Metric',
       y = 'Relative Importance')

@@

```{r site-photo-high, dpi = 300, fig.cap = 'Photo of CHaMP site, CBW05583-415923, with predicted high capacity.'}
knitr::include_graphics('../figures/SitePhoto_CBW05583-415923.jpg')
```

```{r site-photo-low, dpi = 300, fig.cap = 'Photo of CHaMP site, CBW05583-109994, with predicted low capacity.'}
knitr::include_graphics('../figures/SitePhoto_CBW05583-109994.jpg')
```
